---
title: "In Class Exercise 08"
author: "Dew Stella Chan"
date: "October 21, 2024"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse,
               knitr, kableExtra)
```

```{r}
mdata <- read_rds("data/mdata.rds")
```

## Data Sampling
```{r}
set.seed(1234)
HDB_sample <- mdata %>%
  sample_n(100) # set small for faster computating time, for actual modelling, it should be larger
```

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

```{r}
HDB_sample <- HDB_sample %>%
  st_jitter(amount = 5)
```

```{r}
set.seed(1234)
resale_split <- initial_split(HDB_sample, 
                              prop = 6.67/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```


## Multicollinearity check
In order to avoid multicollineariy. In the code chunk below, ggcorrmat() of ggstatsplot is used to plot a correlation matrix to check if there are pairs of highly correlated independent variables.

```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
ggstatsplot::ggcorrmat(mdata_nogeo[, 2:17])

```

## Building a non-spatial multiple linear regression
```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
olsrr::ols_regress(price_mlr)
```

```{r}
vif <- performance::check_collinearity(price_mlr)
kable(vif, 
      caption = "Variance Inflation Factor (VIF) Results") %>%
  kable_styling(font_size = 18) 
```

## Predictive Modelling with gwr
```{r}
bw_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```


```{r}
gwr_adaptive <- gwr.basic(formula = resale_price ~
                            floor_area_sqm + storey_order +
                            remaining_lease_mths + PROX_CBD + 
                            PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL + 
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                            WITHIN_1KM_PRISCH,
                          data=train_data,
                          bw=bw_adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE,
                          longlat = FALSE)
```


```{r}
gwr_adaptive
```
Predicting with test data
```{r}
gwr_bw_test_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=test_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

```{r}

gwr_pred <- gwr.predict(formula = resale_price ~
                          floor_area_sqm + storey_order +
                          remaining_lease_mths + PROX_CBD + 
                          PROX_ELDERLYCARE + PROX_HAWKER + 
                          PROX_MRT + PROX_PARK + PROX_MALL + 
                          PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                          WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
                          WITHIN_1KM_PRISCH, 
                        data=train_data, 
                        predictdata = test_data, 
                        bw=bw_adaptive, 
                        kernel = 'gaussian', 
                        adaptive=TRUE, 
                        longlat = FALSE)
```

## Predictive Modelling: RF method

```{r}
coords <- st_coordinates(HDB_sample)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

```{r}
train_data_nogeom <- train_data %>%
  st_drop_geometry()
```

Calibrating RF Model

```{r}
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data_nogeom)
rf
```
## Predictive Modelling: SpatialML method

```{r}
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + 
                       storey_order + remaining_lease_mths + 
                       PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER + 
                       PROX_MRT + PROX_PARK + PROX_MALL + 
                       PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data_nogeom, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```


```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```


In the code chunk below, predict.grf() of spatialML for predicting re-sale prices in the test data set (i.e. test_data_nogeom)

```{r}
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data_nogeom, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)

```

Next, the code chunk below is used to convert the output from predict.grf() into a data.frame.

```{r}
GRF_pred_df <- as.data.frame(gwRF_pred)
```

Next, the code chunk below is used to convert the output from predict.grf() into a data.frame.

```{r}
test_data_pred <- cbind(test_data, 
                        GRF_pred_df)
```


### Visualising the predicted values
```{r}
ggplot(data = test_data_pred,
       aes(x = gwRF_pred,
           y = resale_price)) +
  geom_point()
```