---
title: "Hands-on Exercise 10"
author: "Dew Stella Chan"

execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

# Overview

```{r}
pacman::p_load(tmap, sf, DT, stplanr, tidyverse,sp, performance, reshape2, ggpubr, tidyverse)
```

## Preparing the Flow Data
```{r}
odbus <- read_csv("data/aspatial/origin_destination_bus_202210.csv")
glimpse(odbus)

```

```{r}
odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE) 
```

```{r}
odbus6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
datatable(odbus6_9)
```


```{r}
write_rds(odbus6_9, "data/rds/odbus6_9.rds")
```

```{r}
odbus6_9 <- read_rds("data/rds/odbus6_9.rds")
```

## Working with Geospatial Data
```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```


```{r}
mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz
```

```{r}
mpsz <- write_rds(mpsz, "data/rds/mpsz.rds")
```


```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```
```{r}
datatable(busstop_mpsz)
```

```{r}
od_data <- left_join(odbus6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```
```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```


```{r}
od_data <- unique(od_data)
```

```{r}
od_data <- left_join(od_data , busstop_mpsz,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 
```

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- unique(od_data)
```


```{r}
od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na() %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarise(MORNING_PEAK = sum(TRIPS))
```

```{r}
write_rds(od_data, "data/rds/od_data_fii.rds")
```

```{r}
od_data_fii <- read_rds("data/rds/od_data.rds")
```

## Visualising Spatial Interaction

```{r}
od_data_fij <- od_data[od_data$ORIGIN_SZ!=od_data$DESTIN_SZ,]
```

```{r}
write_rds(od_data_fij, "data/rds/od_data_fij.rds")
```

```{r}
od_data_fij <- read_rds("data/rds/od_data_fij.rds")
```

## Creating desire lines


```{r}
flowLine <- od2line(flow = od_data_fij, 
                    zones = mpsz,
                    zone_code = "SUBZONE_C")
```

```{r}
write_rds(flowLine, "data/rds/flowLine.rds")
```


```{r}
flowLine <- read_rds("data/rds/flowLine.rds")
```


```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```


```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
  filter(MORNING_PEAK >= 5000) %>%
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```


#Hands Exercise 10b


### Computing Distance Matrix

```{r}
mpsz <- read_rds("data/rds/mpsz.rds")
mpsz
```

### Converting from sf data.table to SpatialPolygonsDataFrame

```{r}
mpsz_sp <- as(mpsz, "Spatial")
mpsz_sp
```

```{r}
dist <- spDists(mpsz_sp, 
                longlat = FALSE)
```


```{r}
head(dist, n=c(10, 10))
```


```{r}
sz_names <- mpsz$SUBZONE_C
```

```{r}
colnames(dist) <- paste0(sz_names)
rownames(dist) <- paste0(sz_names)
```

```{r}
distPair <- melt(dist) %>%
  rename(dist = value)
head(distPair, 10)
```
16.5.5 Updating intra-zonal distances


```{r}
distPair %>%
  filter(dist > 0) %>%
  summary()
```
```{r}
distPair$dist <- ifelse(distPair$dist == 0,
                        50, distPair$dist)
```

```{r}
distPair %>%
  summary()
```

```{r}
distPair <- distPair %>%
  rename(orig = Var1,
         dest = Var2)
```


```{r}
write_rds(distPair, "data/rds/distPair.rds") 
```


```{r}
distPair <- read_rds("data/rds/distPair.rds")
```


## Preparing flow data

```{r}
od_data_fii <- read_rds("data/rds/od_data_fii.rds")
```


```{r}
flow_data <- od_data_fii %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>% 
  summarize(TRIPS = sum(MORNING_PEAK))
```

```{r}
head(flow_data, 10)
```


```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0, flow_data$TRIPS)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0.000001, 1)
```


```{r}
flow_data$ORIGIN_SZ <- as.factor(flow_data$ORIGIN_SZ)
flow_data$DESTIN_SZ <- as.factor(flow_data$DESTIN_SZ)
```

```{r}
flow_data1 <- flow_data %>%
  left_join (distPair,
             by = c("ORIGIN_SZ" = "orig",
                    "DESTIN_SZ" = "dest"))
```


## Preparing Origin and Destination Attributes

```{r}
pop <- read_csv("data/aspatial/pop.csv")
```
 Geospatial data wrangling
```{r}
pop <- pop %>%
  left_join(mpsz,
            by = c("PA" = "PLN_AREA_N",
                   "SZ" = "SUBZONE_N")) %>%
  select(1:6) %>%
  rename(SZ_NAME = SZ,
         SZ = SUBZONE_C)
```

```{r}

flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(ORIGIN_SZ = "SZ")) %>%
  rename(ORIGIN_AGE7_12 = AGE7_12,
         ORIGIN_AGE13_24 = AGE13_24,
         ORIGIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))

```


```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(DESTIN_SZ = "SZ")) %>%
  rename(DESTIN_AGE7_12 = AGE7_12,
         DESTIN_AGE13_24 = AGE13_24,
         DESTIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

### Calibrating Spatial Interaction Models






