---
title: "Hands-on Exercise 2"
author: "Dew Stella Chan"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

The specific questions we would like to answer for this exercise are:

-   are the childcare centres in Singapore randomly distributed throughout the country?

-   if the answer is not, then the next logical question is where are the locations with higher concentration of childcare centres?

### 2.1 Installing and loading of packages

```{r}
pacman::p_load(sf, sp, raster, spatstat, tmap, tidyverse)
```

### 2.2 Data used for this exercise:

-   CHILDCARE downloaded from Data.gov.sg and is in geojson format.

-   MP14_SUBZONE_WEB_PL, downloaded from Data.gov.sg.

-   CostalOutline, provided by SLA.

### 2.3 Spatial Data Wrangling:

Reading of data

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)

sg_sf <- st_read(dsn = "data", layer="CostalOutline")%>%
    st_transform( crs = 3414)

mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")%>%
    st_transform( crs = 3414)

```

#checking for invalid gemetries (when plotting in view)

```{r}
# Check for invalid geometries
invalid_geometries <- st_is_valid(mpsz_sf, reason = TRUE)

# print(invalid_geometries)

# Fix invalid geometries
mpsz_sf_valid <- st_make_valid(mpsz_sf)

```

Plotting of Map:

```{r}
tmap_mode("plot")
tm_shape(mpsz_sf) +
  tm_polygons()+
  tm_shape (childcare_sf) + 
  tm_dots(col = "black", size = 0.01)
```

# Overlaying the map layer and childcare layer

```{r}
tmap_mode("view")
tm_shape(mpsz_sf_valid) +
  tm_polygons()+
  tm_shape (childcare_sf) + 
  tm_dots(col = "black", size = 0.01)
```

## 2.4 GeoSpatial Data Wrangling

### 2.4.1 conversion of sf data t o sp's Spatial\*class

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

View conversion dat using show()

```{r}
show(childcare)
```

```{r}
show(mpsz)
```

```{r}
show(sg)
```

### 4.5.2 Converting the Spatial\* class into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

### To Display the required information

```{r}
print(childcare_sp)
print(sg_sp )
```

## 4.5.3 Converting the generic sp format into spatstat’s ppp format, .

```{r}
childcare_ppp <- as.ppp(childcare_sf)
childcare_ppp
```

```{r}
plot(childcare_ppp)
```

##4.5.4 Handling duplicated points

```{r}
any(duplicated(childcare_ppp))
```

#the count the number of points with duplicates:

```{r}
multiplicity(childcare_ppp)

sum(multiplicity(childcare_ppp)>1)

```

```{r}
sum(multiplicity(childcare_ppp)>1)
```

#The points with darker colour shows are duplicate values.

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

## plotting map to statistic plot

```{r}
tmap_mode('plot')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

# handling duplicates using jitter method.

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

# check for duplicate

```{r}
any(duplicated(childcare_ppp_jit))
sum(multiplicity(childcare_ppp_jit)>1)
```

4.5.5 Creating owin object

```{r}
sg_owin <- as.owin(sg_sf)
plot(sg_owin)
summary(sg_owin)
```

Combining points events objects and owin object:

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
summary(childcareSG_ppp)
plot(childcareSG_ppp)
```

#4.6 First-order Spatial Point Patterns Analysis

```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_childcareSG_bw)
```

4.6.1.2 Rescalling KDE values

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```

4.6.1.2 Rescalling KDE values

```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 1000, "km")
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```

```{r}
 bw.CvL(childcareSG_ppp.km)
```

```{r}
bw.scott(childcareSG_ppp.km)
```

```{r}
bw.ppl(childcareSG_ppp.km)
```

```{r}
bw.diggle(childcareSG_ppp.km)
```

```{r}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

4.6.3 Working with different kernel methods

```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

4.7 Fixed and Adaptive KDE

```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```

4.7.3 Converting KDE output into grid object.(not working)

```{r}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame..im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```

Alternative code:

```{r}
library(raster)
# Convert to RasterLayer and then to SpatialGridDataFrame
raster_kde <- raster(kde_childcareSG.bw)
gridded_kde_childcareSG_bw <- as(raster_kde, "SpatialGridDataFrame")
spplot(gridded_kde_childcareSG_bw)
```

4.7.3.1 Converting gridded output into raster

```{r}
kde_childcareSG_bw_raster <- raster(kde_childcareSG.bw)
```

# coverting gridded oputput tyo

```{r}
kde_childcareSG_bw_raster
```

```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("layer", palette = "viridis") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

pg <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_sf %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_sf %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_sf %>%
  filter(PLN_AREA_N == "JURONG WEST")
```

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol",max.plot = 15)
```

```{r}
par(mfrow=c(2,2))
plot(tm, main = "Tampines",max.plot = 15)
```

```{r}
par(mfrow=c(2,2))
plot(ck, main = "Choa chu Kang",max.plot = 15)
```

```{r}
par(mfrow=c(2,2))
plot(jw, main = "Jurong West",max.plot = 15)
```

```{r}
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)
```

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

```{r}
childcare_pg_ppp.km = rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale.ppp(childcare_jw_ppp, 1000, "km")
```

```{r}
dev.new(width=20, height=20) 
par(mfrow=c(2,2), mar=c(1, 1, 1, 1) + 0.1)

plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

```{r}
dev.new(width=20, height=20) 
par(mfrow=c(2,2), mar=c(1, 1, 1, 1) + 0.1)

plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```

```{r}
dev.new(width=20, height=20) 
par(mfrow=c(2,2), mar=c(1, 1, 1, 1) + 0.1)

plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

4.8 Nearest Neighbour Analysis

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

Based on the outcome of the test,it is statistically significant the childcare in singapore are clustered Specified locations.

4.8.3 Clark and Evans Test: Tampines planning area

```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```


# 5.6 Second-order Spatial Point Patterns Analysis
The steps listed in point 5.1 to 5.5 are the same as point 1st order. 

Refer to section 2.2 and 2.3 for the details of packages and data used in this exercise and the map


5.7.1 Choa Chu Kang planning area
```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```
```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

```{r}
plot(G_CK.csr)
```
5.8.3 Tampines planning area

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```
```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```
```{r}
plot(F_tm.csr)
```

5.9.1 Choa Chu Kang planning area
5.9.1.1 Computing K-fucntion estimate

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```
```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```
5.9.2 Tampines planning area
5.9.2.1 Computing K-fucntion estimation

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```
```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

5.10 Analysing Spatial Point Process Using L-Function
```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```
5.10.1.2 Performing Complete Spatial Randomness Test
5.10.1 Choa Chu Kang planning area

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```
```{r}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

5.10.2 Tampines planning area

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```
```{r}

plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
````


