---
title: "Hands-on Exercise 2"
author: "Dew Stella Chan"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

The specific questions we would like to answer for this exercise are:

-   are the childcare centres in Singapore randomly distributed throughout the country?

-   if the answer is not, then the next logical question is where are the locations with higher concentration of childcare centres?

This webpage consists of two portions from hands-on namely:

1.  First-order spatial analysis

2.  2nd-order spatial analysis

### 2.1 Installing and loading of packages

```{r}
pacman::p_load(sf, sp, raster, spatstat, tmap, tidyverse)
set.seed(1234) #set seed to ensure the result is consistant
```

```{r}
#| eval: false
install.packages("maptools", repos="https://packagemanager.posit.co/cran/2023-10-13")
```

### 2.2 Data used for this exercise:

-   CHILDCARE downloaded from Data.gov.sg and is in geojson format.

-   MP14_SUBZONE_WEB_PL, downloaded from Data.gov.sg.

-   CostalOutline, provided by SLA.

### 2.3 Spatial Data Wrangling:

Reading of data

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)

sg_sf <- st_read(dsn = "data", layer="CostalOutline")%>%
    st_transform( crs = 3414)

mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")%>%
    st_transform( crs = 3414)

```

#checking for invalid geometries (when plotting in view)

```{r}
# Check for invalid geometries
invalid_geometries <- st_is_valid(mpsz_sf, reason = TRUE)

# print(invalid_geometries)

# Fix invalid geometries
mpsz_sf_valid <- st_make_valid(mpsz_sf)

```

Plotting of Map:

```{r}
tmap_mode("plot")
tm_shape(mpsz_sf) +
  tm_polygons()+
  tm_shape (childcare_sf) + 
  tm_dots(col = "black", size = 0.01)
```

## 2.4 Overlay subzone with childcare layer

```{r}
tmap_mode("plot")
tm_shape(mpsz_sf_valid) +
  tm_polygons()+
  tm_shape (childcare_sf) + 
  tm_dots(col = "black", size = 0.01)
```

## 2.5 GeoSpatial Data Wrangling

### 2.5.1 conversion of sf data t o sp's Spatial\*class

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

Verify conversion data using show()

```{r}
show(childcare)
```

```{r}
show(mpsz)
```

```{r}
show(sg)
```

### 2.5.2 Converting the Spatial\* class into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

To Display the required information

```{r}
print(childcare_sp)
print(sg_sp )
```

### 2.5.3 Converting the generic sp format into spatstat’s ppp format

```{r}
childcare_ppp <- as.ppp(childcare_sf)
childcare_ppp
```

```{r}
plot(childcare_ppp)
```

### 2.5.4 Handling duplicated points

```{r}
any(duplicated(childcare_ppp))
```

Codes for counting the number of points with duplicate

No duplicates.

```{r}
tmap_mode('plot')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

### 2.5.5 Plotting map to statistic plot

```{r}
tmap_mode('plot')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

#### 2.5.6 Handling duplicates using jitter method.

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=10, 
                             drop=TRUE)
```

### 2.5.7 Check for duplicate

```{r}
any(duplicated(childcare_ppp_jit))
sum(multiplicity(childcare_ppp_jit)>1)
```

### 2.5.5 owin object

Creating owin object

```{r}
sg_owin <- as.owin(sg_sf)
plot(sg_owin)
summary(sg_owin)
```

Combining points events objects and owin object:

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
summary(childcareSG_ppp)
plot(childcareSG_ppp)
```

### 2.6 First-order Spatial Point Patterns Analysis

```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_childcareSG_bw)
```

#### 2.6.1.2 Rescalling KDE values

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```

#### 2.6.1.2 Rescalling KDE values

```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 1000, "km")
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```

```{r}
 bw.CvL(childcareSG_ppp.km)
```

```{r}
bw.scott(childcareSG_ppp.km)
```

```{r}
bw.ppl(childcareSG_ppp.km)
```

```{r}
bw.diggle(childcareSG_ppp.km)
```

```{r}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

#### 2.6.3 Working with different kernel methods

```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

### 2.7 Fixed and Adaptive KDE\*\*

#### 2.7.1 Computing KDE by using fixed bandwidth\*\*

```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```

#### 2.7.3 Converting KDE output into grid object.

```{r}
gridded_kde_childcareSG_SG<- as.(kde_childcareSG_adaptive,"spatialGridDataFrame")
spplot(gridded_kde_childcareSG_bw)
```

Alternative codes, different from the code listed for the exercise.

```{r}
library(raster)
# Convert to RasterLayer and then to SpatialGridDataFrame
raster_kde <- raster(kde_childcareSG.bw)
gridded_kde_childcareSG_bw <- as(raster_kde, "SpatialGridDataFrame")
spplot(gridded_kde_childcareSG_bw)
```

#### 2.7.3.1 Converting gridded output into raster

```{r}
kde_childcareSG_bw_raster <- raster(kde_childcareSG.bw)
```

```{r}
kde_childcareSG_bw_raster
```

#### 2.7.3.2 Assigning projection system

```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

### 2.7.5 Comparing Spatial Point Patterns using KD\*\*

```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("layer", palette = "viridis") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

pg <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_sf %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_sf %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_sf %>%
  filter(PLN_AREA_N == "JURONG WEST")
```

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol",max.plot = 15)
```

```{r}
par(mfrow=c(2,2))
plot(tm, main = "Tampines",max.plot = 15)
```

```{r}
par(mfrow=c(2,2))
plot(ck, main = "Choa chu Kang",max.plot = 15)
```

```{r}
par(mfrow=c(2,2))
plot(jw, main = "Jurong West",max.plot = 15)
```

```{r}
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)
```

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

```{r}
childcare_pg_ppp.km = rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale.ppp(childcare_jw_ppp, 1000, "km")
```

```{r}
dev.new(width=20, height=20) 
par(mfrow=c(2,2), mar=c(1, 1, 1, 1) + 0.1)

plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

```{r}
dev.new(width=20, height=20) 
par(mfrow=c(2,2), mar=c(1, 1, 1, 1) + 0.1)

plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```

```{r}
dev.new(width=20, height=20) 
par(mfrow=c(2,2), mar=c(1, 1, 1, 1) + 0.1)

plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

## 2.8 Nearest Neighbour Analysis

### 2.8.1 Testing spatial point patterns using Clark and Evans Test\*\*

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

Based on the outcome of the test,it is statistically significant the childcare in singapore are clustered Specified locations.

### 2.8.2 Clark and Evans Test: Choa Chu Kang planning area

```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=99)
```

### 2.8.3 Clark and Evans Test: Tampines planning area

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=99)
```

# 2.9 Second-order Spatial Point Patterns Analysis

The steps listed in point 5.1 to 5.5 are the same as point 1st order.

Refer to section 2.2 to 2.6 for the details of packages and data used in this exercise and the map

### 2.9.1 Computing G-function estimation

#### 2.9.1.1 Choa Chu Kang planning area

The code chunk below is used to compute G-function using *Gest()* of **spatat** package.

```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

#### 2.9.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with G-function

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 99)
```

```{r}
plot(G_CK.csr)
```

#### 2.9.1.2 Tampines planning area

#### 2.9.2.1 Computing G-function estimation

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 99)
```

```{r}
plot(F_tm.csr)
```

## 2.9.2 Analysing Spatial Point Process Using F-Function\*\*

The F function estimates the empty space function F(r) or its hazard rate h(r) from a point pattern in a window of arbitrary shape. In this section, you will learn how to compute F-function estimation by using [*Fest()*](https://rdrr.io/cran/spatstat/man/Fest.html) of **spatstat** package. You will also learn how to perform monta carlo simulation test using [*envelope()*](https://rdrr.io/cran/spatstat/man/envelope.html) of **spatstat** package.

#### 2.9.2.1 Choa Chu Kang Planning Area

#### 2.9.2.1.1 Computing F-function estimation

The code chunk below is used to compute F-function using *Fest()* of **spatat** package.

```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

####2.9.2.1.2 Performing Complete Spatial Randomness Test

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 99)
```

```{r}
plot(F_CK.csr)
```

#### 2.9.2.2 Tampines planning area

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 99)
```

```{r}
plot(F_tm.csr)
```

## 2.9.3 Analysing Spatial Point Process Using K-Function

K-function measures the number of events found up to a given distance of any particular event. In this section, you will learn how to compute K-function estimates by using Kest() of spatstat package. You will also learn how to perform monta carlo simulation test using envelope() of spatstat package.

#### 2.9.3.1 Choa Chu Kang planning area

#2.9.1.1 Computing K-fucntion estimate K-function measures the number of events found up to a given distance of any particular event. In this section, you will learn how to compute K-function estimates by using Kest() of spatstat package. You will also learn how to perform monta carlo simulation test using envelope() of spatstat package.

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

#### 2.9.2.1.2 Performing Complete Spatial Randomness Test\*\*

To confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### 2.9.2.3Analysing Spatial Point Process Using K-Function

### 2.9.2.2 Tampines Planning Area

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

### 2.9.4 Analysing Spatial Point Process Using L-Function

to compute L-function estimation by using Lest() of spatstat package. You will also learn how to perform monta carlo simulation test using envelope() of spatstat package.

### 2.9.4.1 Choa Chu Kang planning area

#### 2.9.4.1.1 computation of L function.

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

#### 2.9.4.1.2 Performing Complete Spatial Randomness Test

o confirm the observed spatial patterns above, a hypothesis test will be conducted. The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value if smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

#### 2.9.4.1.1 Tampines Planning Area

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```
