---
title: "In Class Exercise 02"
author: "Dew Stella Chan"
date: "September 02, 2024"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

```{r}
pacman::p_load(sf, sp, raster, spatstat, tmap, tidyverse)
```



## Issue 2: Creating coastal outline

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")%>%
    st_transform( crs = 3414)
```

```{r}
sg_sf <- mpsz_sf %>%
  st_union()
```


```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```


```{r}
childcare_ppp <- as.ppp(childcare_sf)
plot(childcare_ppp)
```
```{r}
summary(childcare_ppp)
```
Creating owin object from sf data.frame

```{r}
sg_owin <- as.owin(sg_sf)
plot(sg_owin)
```

```{r}
summary(sg_owin)
```

Combining point events object and owin object


```{r}

childcareSG_ppp = childcare_ppp[sg_owin]
```

```{r}
plot(childcareSG_ppp)
```
Kernel Density Estimation of Spatial Point Event

```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 
                                  1000, 
                                  "km")

kde_childcareSG_adaptive <- adaptive.density(
  childcareSG_ppp.km, 
  method="kernel")
plot(kde_childcareSG_adaptive)
```


```{r}
par(bg = '#E4D5C9')

gridded_kde_childcareSG_ad <- maptools::as.SpatialGridDataFrame.im(
  kde_childcareSG_adaptive)
spplot(gridded_kde_childcareSG_ad)
```

```{r}
gridded_kde_childcareSG_ad <- as(
  kde_childcareSG_adaptive,
  "SpatialGridDataFrame")
spplot(gridded_kde_childcareSG_ad)
```

```{r}
tm_shape(gridded_kde_childcareSG_ad) + 
  tm_raster(palette = "viridis") +
  tm_layout(legend.position = c("right", "bottom"), 
            frame = FALSE,
            bg.color = "#E4D5C9")
```
Extracting study area using sf objects

```{r}
pg_owin <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL") %>%
  as.owin()

childcare_pg = childcare_ppp[pg_owin]

plot(childcare_pg)  
```

Monte Carlo Simulation
```{r}
set.seed(1234)
```

##Geospatial Analytics for Social Good: Thailand Road Accident Case Study

Background
Road traffic injuries, WHO.
Road traffic deaths and injuries in Thailand

Importing Traffic Accident Data

```{r}
rdacc_sf <- read_csv("data/data/thai_road_accident_2019_2022.csv") %>%
  filter(!is.na(longitude) & longitude != "", 
         !is.na(latitude) & latitude != "") %>%
  st_as_sf(coords = c(
    "longitude", "latitude"),
    crs=4326) %>%
  st_transform(crs = 32647) 
```

```{r}
# Set tmap mode to "plot" 
tmap_mode("plot") 
tm_shape(rdacc_sf) + tm_dots() 
```
